Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-05 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: removed waveform name comment

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our beyond-GR parameters for specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail these results in Sec.~\ref{sec:results}.
%
\hs{Alessandra, Abhirup, should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are the same (a property known as \emph{isospectrality}) and therefore which
QNM we use to model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode (if excited during the merger) which is
the most likely to appear in the signal.

>>>>

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our beyond-GR parameters for specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail these results in Sec.~\ref{sec:results}.
%
% \hs{Alessandra, Abhirup, should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are isospectral and hence which QNM we use to
model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode (if excited during the merger) which is
the most likely to appear in the signal.
